---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
library(dplyr)
library(tidyr)
library(readr)

deaths <- av %>%
  pivot_longer(cols = starts_with("Death"),
               names_to = "Time",
               values_to = "Death") %>%
  mutate(Time = parse_number(Time),
         Death = tolower(Death)) %>%
  replace_na(list(Death = ""))


head(deaths)
```

Similarly, deal with the returns of characters.

```{r}
returns <- av %>%
  pivot_longer(cols = starts_with("Return"), names_to = "Time", values_to = "Return") %>%
  mutate(Time = parse_number(Time),
         Return = factor(Return, levels = c("yes", "no", "")))

head(returns)
```

Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
avg_deaths <- deaths %>%
  mutate(DeathCount = ifelse(Death == "yes", 1, 0)) %>%  # Convert "yes" to 1, others to 0
  group_by(Name.Alias) %>%                               # Group by Avenger name
  summarise(AvgDeaths = sum(DeathCount)) %>%             # Sum Deaths per Avenger
  summarise(AverageDeaths = mean(AvgDeaths, na.rm = TRUE))  # Calculate overall average

avg_deaths
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

"Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team." (Adam Zhu)

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement
```{r}
# Filter to find Avengers who died at least once
avengers_with_deaths <- deaths %>%
  filter(Death == "yes") %>%
  distinct(Name.Alias) %>%    # Select unique Avengers who died at least once
  count() %>%                 # Count the number of unique Avengers who died
  pull(n)

# Total number of Avengers
total_avengers <- av %>%
  distinct(Name.Alias) %>%
  count() %>%
  pull(n)

# Display results
list(
  Total_Avengers = total_avengers,
  Avengers_With_Deaths = avengers_with_deaths
)

# Adam Zhu
```

### Include your answer
Based on the analysis, the statement provided was incorrect. Here’s the corrected statement: “Out of 173 listed Avengers, my analysis found that 64 had died at least one time after they joined the team.”
This result indicates that 64 Avengers, not 69, experienced at least one death event after joining the team. (Adam Zhu)
